---
title: "Lab 8 - Linear Regression in R"
author: "Robert Heim"
date: "11/21/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#1 Load the tidyverse

```{r include= FALSE}

library(tidyverse)

```

#2 Read in the salmon dataset

```{r message=FALSE}

salmon <- read_csv("SalmonExports.csv")

```

#3 Exploratory Scatterplot 

```{r}

salmon_plot <- ggplot(salmon, aes(x = year, y = salmon_fresh)) +
  geom_point()

salmon_plot
  
# Describing the relationship between year and fresh salmon imports using linear regression makes sense becaue the data appears linear 

```

#4 Perform a linear regression

```{r}

# Perform linear regression to describe the trend in total volume of fresh salmon imported to the U.S.from 1990 - 2017.

salmon_model <- lm(salmon_fresh ~ year, data = salmon)

salmon_model

# Determine the model equaiton: recall... y= B0 + B1x + E

# Model equation: Salmon Imports(tons) = - 14982940 + 7550(year)

# The slope means that each year, salmon imports increase by 7550 pounds on average.
# The y-intercept means that at year 0, salmon imports are expected to be -14982940 pounds. This y-intercept is concerning because imports cannot be a negative number. This model should not be used to extrapolate for past values of salmon imports.

```

#6 Model diagnostics

Use plot(model_name) to view model diagnostics in the ‘Plots’ tab (press Enter in the Console to contine to subsequent diagnostic plots). 

```{r}

plot(salmon_model)

# The residuals appear normally distributed

# Points 15, 24, 25, and 28 appear to be outliers

# Use the code below to view all four diagnostic plots at once, and have them appear in a knitted document. 

par(mfrow = c(2,2))
plot(salmon_model)

```

# 7 Explore model fit and significance

Use summary(model_name) to see the detailed model information, including model fit information (e.g. R2 and coefficient standard errors) and statistics.

```{r}

summary(salmon_model) 

# Year significantly predicts salmon imports (p < 0.001)

# The R-squared value means that 94.8% of variance in salmon imports is explained by year. 

# 5.2% of the variance in salmon imports in not explained by year. 

```

# 8 Prepare a final graph of the linear regression model with a 95% confidence interval

```{r}

salmon_final_graph <- ggplot(salmon, aes(x = year, y = salmon_fresh)) +
geom_point() +
geom_smooth(method = lm, se = TRUE, size = 0.5, color = "gray20") +
theme_bw() +
scale_x_continuous(expand = c(0,0), limits = c(1990, 2017)) +
labs(x = "Year", y = "U.S. Annual Fresh Salmon Imports\n(1,000 pounds)")

salmon_final_graph


# geom_smooth() adds confidence bands 

```

# Fig. 1 Total U.S. fresh salmon imports as a function of year between 1990-2017. Year predicted a signficant proportion of variance in salmon imports (R-squared= 0.948, F(1,26) = 471.9, p<0.001). 

#9 Make predicitons using the model

Let’s say we want to use our model to make predictions for U.S. fresh salmon imports in years 2022, 2024, 2026 and 2028.


```{r}

# Create a new data frame called ‘new_years’ containing the sequence of years that we want to make predictions for. The column name we create is the same as the variable name ‘year’ that the model uses.

new_years <- data.frame(year = c(2022, 2024, 2026, 2028))

# Enter the new_years data frame into the predict() function. Include interval = “confidence” to report the lower and upper bounds of the 95% CI for model fit at each year.

future_predict <- predict(salmon_model, newdata = new_years, interval = "confidence")

future_predict

# The prediction outcomes are in the fit column. 
# Bind the prediction outcomes with the new_years data frame to create a useful table of predicted imports with lower and upper limits. 

predictions <- data.frame(new_years, future_predict)

predictions


```

# 10 Find Pearson’s r (correlation)

Use the cor.test() function to find Pearson’s r for the linear relationship described between year and salmon imports.

```{r}

# Explore the correlation between year and salmon imports

salmon_correlation <- cor.test(salmon$year, salmon$salmon_fresh)

salmon_correlation

# There is a significant and strong positive correlation between year and salmon imports. 

```

# Concluding statement about the relationship

Year signficantly predicts U.S. total salmon imports (b = 7550, t(26)= 21.7, p<0.001) with a strong positive correlation between the two (Pearson's r = 0.974). The overal model (imports = - 14982940 + 7550(year); in 1,000 pounds) explains a signficant amount of variance in salmon imports (F(1,26) = 471.9, p<0.001, R2= 0.95). 




